Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-08-13 paper/main.tex
update scripts

% brain mapping via fMRI
Studies in the field of neuropsychology and neuroimaging
(\citep[e.g.,][]{penfield1950cerebral, fox1984noninvasive}) have shown that different
parts of the brain are specialized for different perceptual and cognitive
functions.
% occipital cortex for vision -> two pathways
The occipital cortex is considered to be primarily involved in the early stages
of visual perception and giving rise to two distinct pathways that process
different kind of visual information:
% dorsal vs ventral pathway
a dorsal stream (the ``where pathway'') that leads into the parietal lobe, and a
ventral stream (the ``what pathway'') that leads into the temporal lobe
\citep{goodale1992separate, mishkin1982contribution}.

\todo[inline]{Goodale and Milner (1992), which is literally a call for what and
    how pathways. The function of these pathways remains in dispute but its
    clear that where (spatial processing) must take place in both streams. You
should probably acknowledge that.}
% PPA
A classic example of a higher-level visual area in the ventral pathway is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location != functional location
The PPA is located in the posterior parahippocampal gyrus including adjacent
regions of the fusiform gyrus and anterior lingual gyrus
\citep{epstein2008parahippocampal}.
% neural correlate of scene perception
Increased hemodynamic activity is observed in the PPA when participants view
pictures of landscapes, buildings or landmarks (compared to e.g. pictures of
faces or tools) during blood oxygenation level-dependent functional magnetic
resonance imaging (BOLD fMRI) (\citep[see reviews][]{epstein2014neural, aminoff2013role}).
% \citep{aguirre1998area, epstein2014neural, epstein1998ppa, troiani2012object}.

@@

% av intro
First, we analyzed data from the movie that offered ecologically more valid
visual stimulation than a paradigm using blocks of pictures.
% AV primary contrast
The movie's primary $t$-contrast that compared cuts to a setting that was not
depicted before to cuts within a recurring setting (\texttt{vse\_new >
vpe\_old}) yielded three significant clusters (see Table~\ref{tab:res-av-group1},
Figgure~\ref{fig:group-slices}).
% PPA (others the clusters extend into).
One cluster spans across the midline and comprises parts of the intracalcarine
and cuneal cortex, the lingual gyrus and retrosplenial cortex, the occipital and
temporal fusiform gyrus, and the parahippocampal cortex (reported from posterior
to anterior) in both hemispheres.
% LOC
Two additional bilateral clusters are located in the superior lateral occipital
cortex.

@@

\begin{figure*}[tbp]
\centering
    \includegraphics[width=\linewidth]{figures/subs-thresh-ppa}
    \caption{Fixed-effects individual-level GLM results ($Z$>3.4, $p$<.05
        cluster-corrected).
        Individual brains are aligned via non-linear
        transformation to a study-specific T2* group template that is
        co-registered to the MNI152 template with an affine transformation (12
        degrees of freedom).
        The results of the audio-description's primary
        $t$-contrast (blue) that compares geometry related nouns to non-
        geometry related nouns spoken by the narrator
        (\texttt{geo, groom > all non-geo}) are overlaid over the movie's
        primary $t$-contrast (blue) that compares cuts to setting depicted for
        the first time with cuts within a recurring setting
        (\texttt{vse\_new > vpe\_old}).
        Black:
        outline of participant-specific PPA ROIs.
        Light gray: The
        audio-description's field of view \citep{hanke2014audiomovie}.
        To facilitate comparisons across participants, we chose the same horizontal
        slice (x=-11) for all participants as this slice depicts voxels of
        significant clusters in almost all participants.
        The figure does not show voxels of the left cluster of the movie stimulus
        in sub-09 and sub-18, and voxels of the right cluster of the movie
        stimulus in sub-15.}
    \label{fig:subs-thresh-ppa}
\end{figure*}

@@

\section*{Data (and code) availability}

@@

\todo[inline]{Please also split "Code and Data availability" into two sections}

@@

>>>>

% brain mapping via fMRI
Studies in the field of neuropsychology and neuroimaging
(\citep[e.g.,][]{penfield1950cerebral, fox1984noninvasive}) have shown that
different parts of the brain are specialized for different perceptual and
cognitive functions.
% occipital cortex for vision -> two pathways
the occipital cortex is considered to be primarily involved in the early stages
of visual perception and giving rise to two distinct [but interacting] pathways
that serve different functions:
% dorsal vs ventral pathway
a) a dorsal stream (the ``how pathway'') that [leads into the parietal lobe,]
and supports [nonconscious] visual guidance of action, and
b) a ventral stream (the ``what pathway'') that [leads into the temporal lobe
and] supports [conscious] perception and recognition.
\citep{goodale1992separate, milner2017two, mishkin1982contribution}.

\todo[inline]{reviewer #1: its clear that where (spatial processing) must take
place in both streams. You should probably acknowledge that.}
% PPA
A classic example of a higher-level visual area in the ventral pathway is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location != functional location
The PPA is located in the posterior parahippocampal gyrus including adjacent
regions of the fusiform gyrus and anterior lingual gyrus
\citep{epstein2008parahippocampal}.
% neural correlate of scene perception
Increased hemodynamic activity is observed in the PPA when participants view
pictures of landscapes, buildings or landmarks (compared to e.g. pictures of
faces or tools) during blood oxygenation level-dependent functional magnetic
resonance imaging (BOLD fMRI) (\citep[see reviews][]{epstein2014neural,
aminoff2013role}).
% \citep{aguirre1998area, epstein2014neural, epstein1998ppa, troiani2012object}.

@@

% av intro
First, we analyzed data from the movie that offered ecologically more valid
visual stimulation than a paradigm using blocks of pictures.
% AV primary contrast
The movie's primary $t$-contrast that compared cuts to a setting that was not
depicted before to cuts within a recurring setting (\texttt{vse\_new >
vpe\_old}) yielded three significant clusters (see Table~\ref{tab:res-av-group1},
Figure~\ref{fig:group-slices}).
% PPA (others the clusters extend into).
One cluster spans across the midline and comprises parts of the intracalcarine
and cuneal cortex, the lingual gyrus and retrosplenial cortex, the occipital and
temporal fusiform gyrus, and the parahippocampal cortex (reported from posterior
to anterior) in both hemispheres.
% LOC
Two additional bilateral clusters are located in the superior lateral occipital
cortex.

@@

\begin{figure*}[tbp]
\centering
    \includegraphics[width=\linewidth]{figures/subs-thresh-ppa}
    \caption{Fixed-effects individual-level GLM results ($Z$>3.4, $p$<.05
        cluster-corrected).
        Individual brains are aligned via non-linear
        transformation to a study-specific T2* group template that is
        co-registered to the MNI152 template with an affine transformation (12
        degrees of freedom).
        The results of the audio-description's primary
        $t$-contrast (blue) that compares geometry related nouns to non-
        geometry related nouns spoken by the narrator
        (\texttt{geo, groom > all non-geo}) are overlaid over the movie's
        primary $t$-contrast (red) that compares cuts to setting depicted for
        the first time with cuts within a recurring setting
        (\texttt{vse\_new > vpe\_old}).
        Black:
        outline of participant-specific PPA ROIs reported by
        \citet{sengupta2016extension}.
        Light gray: The
        audio-description's field of view \citep{hanke2014audiomovie}.
        To facilitate comparisons across participants, we chose the same horizontal
        slice (x=-11) for all participants as this slice depicts voxels of
        significant clusters in almost all participants.
        The figure does not show voxels of the left cluster of the movie stimulus
        in sub-09 and sub-18, and voxels of the right cluster of the movie
        stimulus in sub-15.}
    \label{fig:subs-thresh-ppa}
\end{figure*}

@@

\section*{Data availability}

@@

Scripts to generate the results as Datalad \citep{datalad} are published to a
GIN repository \href{https://gin.g-node.org/chaeusler/studyforrest-ppa-analysis}{\url{gin.g-node.org/chaeusler/studyforrest-ppa-analysis}} (DOI assignment pending).

@@

\subsection*{Surface maps}

\todo[inline]{Just mention this part in the main document via 'the corresponding
surface plots (Fig xy) can be found in the supplemental materials'?}
\todo[inline]{write intro regarding purpose of additional surface plots?}

% reconstruction MNI
Cortical reconstruction of the MNI152 T1-weighted brain template
[MNI152\_T1\_1mmi\_brain.nii.gz] was performed using
\href{https://surfer.nmr.mgh.harvard.edu}{FreeSurfer} v7.1.1
\citep{dale1999cortical}.
% reconstruction: individual results
Subject-specific cortical surfaces are part of pre-existing
\href{https://github.com/psychoinformatics-de/studyforrest-data-freesurfer.git}{dataset}
that used FreeSurfer v5.3.0 to reconstruct surfaces from T1-weighted images and
an additional high-resolution T2-weighted image.

% projection
Results were projected onto the corticals surface using
\href{https://surfer.nmr.mgh.harvard.edu}{FreeSurfer's} \citep{dale1999cortical}
'mri\_vol2surf' command.
% smoothing
Group and individual PPAs were additionally patially smoothed applying a
Gaussian kernel with full width at half maximum (FWHM) of \unit[3.0]{mm}.

\subsubsection*{Group results}
Group results see Table~\ref{tab:res-ao-group1} for AO results,
Table~\ref{tab:res-av-group1} for AV results, and
Figure~\ref{fig:grp_ao_av_c1_surf} for surface maps.

\begin{figure*}[tbp]
\centering
    \includegraphics[width=\linewidth]{figures/grp_ao_av_c1_surf}

    \caption{Mixed-effects group-level (N=14) clusters ($Z$>3.4, $p$<.05
      cluster-corrected) of activity correlated with the processing of spatial
      information are projected onto the surface of the MNI152
      T1-weighted template.
      %
      The results of the audio-description's primary $t$-contrast (blue) that
      compares geometry-related nouns spoken by the narrator to non-spatial
      nouns (\texttt{geo, groom} > all non-spatial categories) are overlaid
      on the movie's primary $t$-contrast (red) that compares cuts to a
      setting depicted for the first time with cuts within a recurring setting
      (\texttt{vse\_new > vpe\_old}).
      %
      Black: outline of the union of individual PPA ROIs reported by
      \citet{sengupta2016extension}.
    }
    \label{fig:grp_ao_av_c1_surf}
\end{figure*}

\subsubsection*{Stability}
Group results see Figure~\ref{fig:grp_ao_av_stability} for surface maps.

\begin{figure*}[tbp]
\centering
    \includegraphics[width=\linewidth]{figures/grp_ao_av_stability_surf}
    \caption{Overlap of significant clusters ($Z$>3.4; $p$<.05, cluster
        corrected) across all contrasts for both naturalistic stimuli.
        The audio-description's contrasts 1-8 (blue)
        are overlaid over the audio-visual movie's contrasts 1-5 (red;
        \ref{tab:contrasts}).
        Cluster are projected onto the surface of the
        MNI152 T1-weighted template.
        Black: outline of the union of individual PPA ROIs reported by
        \citet{sengupta2016extension}.
        }
    \label{fig:grp_ao_av_stability}
\end{figure*}

\subsubsection*{Individual resultsStability}

Individual results see Figure~\ref{fig:subjs_ao_av_c1_surf} for surface maps.
\todo[inline]{RSC in of AV is essentially not visible}

\begin{figure*}[tbp]
\centering
    \includegraphics[width=\linewidth]{figures/subjs_ao_av_c1_surf}
    \caption{Fixed-effects individual-level GLM results ($Z$>3.4, $p$<.05
        cluster-corrected)
        projected onto subject-specific brain surface that were
        [re]constructed from individual T1-weighted brain images.
        The results of the audio-description's primary
        $t$-contrast (blue) that compares geometry related nouns to non-
        geometry related nouns spoken by the narrator
        (\texttt{geo, groom > all non-geo}) are overlaid over the movie's
        primary $t$-contrast (red) that compares cuts to setting depicted for
        the first time with cuts within a recurring setting
        (\texttt{vse\_new > vpe\_old}).
        Black: outline of participant-specific PPA ROIs reported by
        \citet{sengupta2016extension}.
        }
    \label{fig:subjs_ao_av_c1_surf}
\end{figure*}
